Jacobi Elliptic Functions in $\RR\times\RR$

### 2016-03-24 number_theory/hw/hw7.tex
number theory

\item Suppose that $p$ is prime and $k$ is a positive integer, and $x\in\ZZ_{p}$.
\begin{enumerate}
\item Prove that $x^{p^{k}}\equiv x\mod p$.

\item Prove that $x^{p^{k}-1}\equiv 1\mod p$ if and only if $\gcd{(x,p)}=1$.
\end{enumerate}

\item Let $p,n$, and $r$ be non-negative integers with $p$ prime. Further, assume that $r$ is the
remainder when $n$ is divided by $p-1$. Prove that $a^{n}\equiv a^{r}\mod p$.

\item Assume that $n$ is a positive integer. Prove that $1^{n}+2^{n}+3^{n}+4^{n}$ is divisible by
5 if and only if $n$ is not divisible by 4.

\item Determine whether there exists a solution to each of the following systems of congruences.  If
there is a solution, find all solutions to the system by writing the solution set as a single residue
class modulo $n$ for some $n\geq 2$. If there is no solution, prove that there is no solution.
\begin{enumerate}
\item $x\equiv5\mod7$\\
$x\equiv0\mod4$
\begin{align*}
x_{0}&=a_{1}c_{1}d_{1}+1_{2}c_{2}d_{2}\\
&= 5c_{1}d_{1} + 0c_{2}d_{2}\\
&= 5\cdot4\cdot d_{1}\\
&= 5\cdot4\cdot2\\
&= 40\\
&\equiv 12\mod28
\end{align*}

>>>>

\item Suppose that $p$ is prime and $k$ is a positive integer, and $x\in\ZZ_{p}$.
\begin{enumerate}
\item Prove that $x^{p^{k}}\equiv x\mod p$.
\begin{quote}
We can rewrite $x^{p^{k}}$ as $(((x^{p})^{p})^{p}...)^{p}$, where there are $k\ p'\text{s}$. We
can mod out by $p$, and we get $(((x^{p})^{p})^{p}...)^{p}$, where there are $k-1\ p'\text{s}$. If
we mod out $k-2$ more times, we get $x^{p}\equiv x\mod p$, which by the corollary to Fermat's
Little Theorem means $x^{p^{k}}\equiv x\mod p$.
\end{quote}
\item Prove that $x^{p^{k}-1}\equiv 1\mod p$ if and only if $\gcd{(x,p)}=1$.
\begin{quote}
$\Rightarrow)$ Assume that $gcd(x,p)\neq1$. Then $x\equiv0\mod p$, since the only thing not coprime
with $p$ is a multiple of $p$. We can raise both sides of the congruence to the ${p^{k}-1}$
power to produce
\[
x^{p^{k}-1}\equiv 0^{p^{k}-1} \equiv 0 \ \cancel{\equiv}\ 1 \mod p.
\]
\noindent\\
$\Leftarrow)$ Suppose $gcd(x,p)=1$. Then $x$ has an inverse in $\ZZ_{p}$. We know by part $(a)$
that $x^{p^{k}}\equiv x\mod p$, so $x^{-1}\cdot x^{p^{k}}\equiv x\cdot x^{-1}\equiv 1\mod p$.
\end{quote}
\end{enumerate}

\item Let $p,n$, and $r$ be non-negative integers with $p$ prime. Further, assume that $r$ is the
remainder when $n$ is divided by $p-1$. Prove that $a^{n}\equiv a^{r}\mod p$.
\begin{quote}
Since $r$ is the remainder, we have that $n = q(p-1)+r$ by the division algorithm. Since
$p,n$, and $r$ are non-negative, we can say $a^{n}=a^{q(p-1) + r}$, which is the same as
$a^{n}=a^{q(p-1)}a^{r}$. By Fermat's Little Theorem, we have that $a^{p-1}=1\mod p$, so
$a^{n}\equiv 1^{q}a^{r}\mod p$, or $a^{n}\equiv a^{r}\mod p$.
\end{quote}

\item Assume that $n$ is a positive integer. Prove that $1^{n}+2^{n}+3^{n}+4^{n}$ is divisible by
5 if and only if $n$ is not divisible by 4.
\begin{quote}
$\Rightarrow)$ Suppose $4\mid n$. Then $n=4k,\ k\in\ZZ$, so we have
\begin{align*}
1^{n}+2^{n}+3^{n}+4^{n}&\equiv 1^{4k}+2^{4k}+3^{4k}+4^{4k}\\
&\equiv 1^{4^k}+2^{4^k}+3^{4^k}+4^{4^k}\\
&\equiv 1^k+1^k+1^k+1^k\mod5\\
&\ \cancel{\equiv}\ 0\mod5.
\end{align*}
\noindent\\
$\Leftarrow)$ Suppose $4\nmid n$. Then $n\ \cancel{\equiv}\ 0\mod4$. Suppose that $n\equiv1\mod4$.
We now have $n=4k+1$ for some $k\in\ZZ$. By $(3)$, we know that $a^{n}\equiv a^{r}\mod p$.
Similarly, we can say $a^{n} + b^{n} + c^{n} + d^{n}\equiv a^{r}+b^{r} + c^{r} + d^{r}\mod p$.
Then, we have $1^{n}+2^{n}+3^{n}+4^{n}\equiv 1 + 2 + 3 + 4\mod5\equiv0\mod5$. Suppose $n\equiv2\mod4$.
Then we have $1^{n}+2^{n}+3^{n}+4^{n}\equiv 1 + 4 + 9 + 16\equiv1-1+1-1\mod5\equiv0\mod5$. Finally,
suppose $n\equiv3\mod4$. Then $1^{n}+2^{n}+3^{n}+4^{n}\equiv1+8+27+64\equiv1+2+3-1\equiv0\mod5$.
\end{quote}

\item Determine whether there exists a solution to each of the following systems of congruences.  If
there is a solution, find all solutions to the system by writing the solution set as a single residue
class modulo $n$ for some $n\geq 2$. If there is no solution, prove that there is no solution.
\begin{enumerate}
\item $x\equiv5\mod7$\\
$x\equiv0\mod4$
\begin{quote}
Since $gcd(7,4)=1$, we can use Chinese remainder theorem, which tells us
\begin{align*}
x_{0}&=a_{1}c_{1}d_{1}+1_{2}c_{2}d_{2}\\
&= 5c_{1}d_{1} + 0c_{2}d_{2}\\
&= 5\cdot4\cdot d_{1}\\
&= 5\cdot4\cdot2\\
&= 40\\
&\equiv 12\mod28
\end{align*}
\end{quote}

### 2016-03-24 pde_two/elliptic.tex
number theory

We will begin by making a comparison of JEFs to Trig functions, with which we are familiar. Consider
$\sin{}$ and $\cos{}$. There are many definitions of these two functions, but we will consider a
definition that relates the two functions to ODEs, namely the system
\[
\begin{cases}
\dot{x}=y\\
\dot{y}=-x
\end{cases}.
\]
Under certain initial conditions, this leads to a basis composed of $\sin{}$ and $\cos{}$. It is with
this analogy that we can consider the JEFs.

Suppose we have a system of ODEs as follows
\[
\begin{cases}
\dot{x}=yz\\
\dot{y}=-zx\\
\dot{z}=-k^{2}xy
\end{cases}.
\]
These may seem like arbitrary selections for the system, but, as we will see, they are not. We need
to define $k$ before continuing; it is known as the modulus. We will also want to define the
complimentary modulus, $\kappa=\sqrt{1-k^{2}}$. If we return to our ODEs, we can define the JEFs
as solutions to the system under the following initial conditions:
\[
\begin{cases}
sn(0,k)=x(0)=0\\
cn(0,k)=y(0)=1\\
dn(0,k)=z(0)=1\\
\end{cases}.
\]

We can draw a relationship between $x$, $y$, and $z$ by forming a basis, much like we did with the
ODE definitions of $\sin{}$ and $\cos{}$. Namely, we have that
\[
\{ x,y,z \}\to\{ sn(t,k),cn(t,k),dn(t,k) \}.
\]
This gives rise to the immediate result that
\[
\begin{cases}
{d\over\dt}sn(t,k)=cn(t,k)dn(t,k)\\
{d\over\dt}cn(t,k)=-dn(t,k)sn(t,k)\\
{d\over\dt}dn(t,k)=-k^{2}sn(t,k)cn(t,k)
\end{cases}.
\]

@@

\begin{minipage}{0.5\textwidth}
\[
\begin{cases}
cn(u,k)=y\\
sn(u,k)={x\over a}\\
dn(u,k)={r\over a}
\end{cases}
\]
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{center}
\includegraphics[scale=0.5]{ellipsis}\\

>>>>

We will begin by making a comparison of JEFs to Trig functions, with which we are familiar. Consider
$\sin{t}$ and $\cos{t}$. There are many definitions of these two functions, but we will consider a
definition that relates the two functions to ODEs, namely the system
\[
\begin{cases}
\dot{x}=y\\
\dot{y}=-x
\end{cases}.
\]
Under certain initial conditions, this leads to a basis composed of $\sin{t}$ and $\cos{t}$. It is with
this analogy that we can consider the JEFs.

Suppose we have a system of ODEs as follows
\[
\begin{cases}
\dot{x}=yz\\
\dot{y}=-zx\\
\dot{z}=-k^{2}xy
\end{cases}.
\]
These may seem like arbitrary selections for the system, but, as we will see, they are not. We need
to define $k$ before continuing; it is known as the modulus. On an ellipse, the value of $k$ is given
by
\[
k=\sqrt{1-{a^{2}\over b^{2}}},
\]
where $a$ is the major axis and $b$ is the semi-major axis.
We can also define $\kappa=\sqrt{1-k^{2}}$, known as the complimentary modulus. If we return to
our system, we can define the JEFs as solutions to the system under the following initial conditions:
\[
\begin{cases}
sn(0,k)=x(0)=0\\
cn(0,k)=y(0)=1\\
dn(0,k)=z(0)=1\\
\end{cases}.
\]

We can draw a relationship between $x$, $y$, and $z$ by forming a basis, much like we did with the
ODE definitions for $\sin{t}$ and $\cos{t}$. Namely, we have that
\[
\{ x,y,z \}\to\{ sn(t,k),cn(t,k),dn(t,k) \}.
\]
This gives rise to the immediate result that
\[
\begin{dcases*}
{d\over\dt}sn(t,k)=cn(t,k)dn(t,k)\\
{d\over\dt}cn(t,k)=-dn(t,k)sn(t,k)\\
{d\over\dt}dn(t,k)=-k^{2}sn(t,k)cn(t,k)
\end{dcases*}.
\]

@@

\begin{minipage}{0.5\textwidth}
\[
\begin{dcases*}
cn(u,k)=y\vphantom{\frac{0}{0}}\\
sn(u,k)={x\over a}\\
dn(u,k)={r\over a}
\end{dcases*}
\]
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{center}
\includegraphics[scale=0.5]{ellipsis}\\
